---
title: "carpentR"
output: github_document
---

This R Package implements a model from Stephen R. Carpenter's 1992 
[1] [paper](https://doi.org/10.1007/978-1-4612-4410-3_23) on 
predicting lake algal blooms using plankton dynamics and other variables, and
was developed specifically for use in Sam Fey's Ecology class at Reed College.

The developmental version of this package can be installed with the following
code:

```{r, eval = FALSE}
# install the remotes package (if you haven't yet)
# install.packages("remotes")

# install the carpentR package!
remotes::install_github("simonpcouch/carpentR")
library(carpentR)
```

Each of the functions exported by this package begin with the prefix `carpenter_*`--typing the first few letters will show all of the different functions available in {carpentR}!

```{r load, include = FALSE, warning = FALSE, message = FALSE}
devtools::load_all()
library(xtable)
```

The primary function from this package is called `carpenter_model`, which runs the simulations detailed in the paper based on the parameters given. An example is shown below!

```{r model}
model_data <- carpenter_model(nanoplankter_diameter = 5.0,
                              alga_diameter = 40,
                              length_herbivore = 1,
                              death_rate_herbivore = .1,
                              p_influx_rate = 0.7,
                              p_outflow_rate = 0.0003,
                              mixed_layer_depth = 2.5)

head(model_data)
```

{carpentR} also provides several functions to interface with this data. You might be interested in using any of the `carpenter_viz_*` functions:

```{r viz}
carpenter_viz_phosphorus(model_data)

carpenter_viz_chlorophyll(model_data)

carpenter_viz_zooplankton(model_data)
```

The `carpenter_visualize()` function will call each of the three visualization functions (in the above order), and you can add aesthetic layers to these plots as you would with any other `ggplot`:

```{r viz2}
carpenter_viz_phosphorus(model_data) + ggplot2::theme_minimal()
```

Finally, the package supplies a function called `carpenter_pivot()` that allows you to pivot between wide and long data. Here's an example:

```{r pivot}
model_data_long <- carpenter_pivot(model_data)

head(model_data_long)
```

Depending on what you want to visualize, either of the default output structure or the "longer" data will be properly tidy (and thus easier to use with `ggplot2` and other packages from the `tidyverse`.)

Happy modeling!

Note that "original source code" is mentioned throughout this package's 
source---this is in reference to an original implementation of the model in 
Pascal.

---------------------------------------------------------------

[1] Carpenter S.R. (1992) Destabilization of Planktonic Ecosystems and Blooms 
of Blue-Green Algae. In: Kitchell J.F. (eds) Food Web Management. Springer 
Series on Environmental Management. Springer, New York, NY